### 2025-01-02 skripsi/chapter4.tex
finish second scenario, fix figures and spelling ch4, start third scenario

% TODO: add figures and tables of generated data and some intermediate steps.
\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/clean_perturbed_solution_0.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_noise_levels}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean_0.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/derivative_clean_0.pgf}
    \end{adjustbox}
    \caption{}\label{fig:derivative}
  \end{subfigure}
  \caption{(\subref{fig:antiderivative_noise_levels}) A generated function with no noise (clean), low noise level (5\%), medium noise level (10\%), and high noise levels (50\%).\ (\subref{fig:antiderivative}) Antiderivative function with no noise.\ (\subref{fig:derivative}) Derivative function with no noise.}
\end{figure}
% TODO: interesting note that model trains better on function to coefficient task when the antiderivative is computed from the randomly generated derivative than if the derivative is computed from the randomly generated antiderivative. THIS NEEDS A REASON AND ANALYSIS SO MIGHT NOT BE NEEDED IN THIS WRITING. MAYBE A GOOD IDEA FOR A SEPARATE PIECE

@@

After this, in the third step, the solution functions \(u \) are generated with 25 modes in time and 25 in space and 40 unique function samples for each viscosity value. In total 120 unique solution functions are generated. To ensure the same solutions are generated every time, the random number generator uses a new instance of the generator with the same seed number. This results in the solutions being the same for all viscosity and other parameters of the equation are adapted to this.

As for the forcing term, it is computed using \lccref{eq:forced_viscous_burgers}. The solution field \(u\) and forcing term \(f\) are modeled with Fourier series in \lccrefs{eq:fourier_field,eq:fourier_force} where \(k\) is a vector of spatial and temporal wave number such that \(k_t\) is the temporal wave number and \(k_x\) is the spatial wave number.
\begin{align}
  u\left(x, t \right) & = \sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:fourier_field} \\
  f\left(x, t \right) & = \sum_{k_{x}}\sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:fourier_force}
\end{align}
Substituting the terms in \lccref{eq:forced_viscous_burgers} with the respective Fourier series, we get the equation \lccref{eq:forced_viscous_burgers_fourier_substitution}. A problem one notices is the nonlinear term \(u\pdv{u}{x}\). A naive approach would multiply all wave numbers terms with each other.
\begin{align}
  u\pdv{u}{x} = \left(\sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}\right) \times \left(\sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})\hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}\right)
\end{align}
This is computationally expensive operation with \(N^2\) multiplications \autocite{lariosMATH934BURGERS2021,robertsDealiasedConvolutionsPseudospectral2011,shenSpectralMethodsAlgorithms2011,orszagComparisonPseudospectralSpectral1972}. To avoid this complexity, the term is first reformulated into \(\pdv{x}(u^2/2)\). Then to avoid aliasing, the coefficients are padded with 50\% zeros such that the padded coefficients are 3/2 times the size of the original \autocite{orszagEliminationAliasingFiniteDifference1971, lariosMATH934BURGERS2021}. Then we use the padded coefficients to compute the function values in the physical domain using the inverse transform. And then, the point wise squaring operation is performed and transform the results back to spectral domain. The resulting coefficients are then trimmed back to their original size before padding. Finally, we multiply the resulting coefficients \(\hat{uu}_k\) by the derivative constants. This operation is much more efficient with only \(1.5 N\) multiplications. This is still more efficient than the naive approach even after taking into account the transforms involved adds \(O(2\times 1.5Nln(1.5N))\) operations if using a Fast Fourier Transform algorithm. Putting all the above together, results in \lccref{eq:forced_viscous_burgers_fourier_substitution}.
\begin{equation}
  \begin{split}
    \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x+k_{t}t)} =
     & \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}) \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}                                                              \\
     & + \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})\hat{uu}_k e^{2\pi i(k_{x}x+k_{t}t)}                                                            \\
     & - \nu\sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})^2\hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:forced_viscous_burgers_fourier_substitution}
  \end{split}
\end{equation}
\begin{align}
  \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k & = \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}) \hat{u}_k + (2\pi i k_{x})\hat{uu}_k - \nu(2\pi i k_{x})^2\hat{u}_k \label{eq:forced_viscous_burgers_fourier} \\
  \hat{f}_k                           & = (2\pi i k_{t}) \hat{u}_k + (2\pi i k_{x})\hat{uu}_k - \nu(2\pi i k_{x})^2\hat{u}_k \label{eq:forced_viscous_burgers_coeff}
\end{align}
After simplifying the equation, we get the coefficients as in \lccref{eq:forced_viscous_burgers_coeff}. Using this equation, the exact forcing term corresponding to the randomly generated solution can be computed with relatively low cost. Finally, the function values of both the solutions and forcing terms are perturbed by adding Gaussian noise with a mean of zero and standard deviation of 5\% of the function value standard deviation to the values of the inverse transform. The perturbed coefficients are then recomputed from the sum and the final perturbed functions are obtained.

An example of the solution and forcing term pairs that has been generated is shown in \lccref{fig:burgers_data}. Using the same seed we can see that the solution functions of the same sample are the same for the most part. However, with the forcing terms, we see much larger amplitudes for higher frequencies for larger viscosity values. This is inline with \lccref{eq:forced_viscous_burgers_coeff}. The squared multiplier gets large much more quickly for higher frequencies due to the squaring operation. The since this effect is controlled by the viscosity via multiplication, the higher the viscosity, the larger the high frequency amplitudes of the forcing term. This in other words mean that more viscous fluids require quadratically more force to affect the solution in the same way with lower viscosities.

@@

\section{Data Retrieval}\label{sec:data_retrieval}
\noindent The second data acquisition approach retrieves data of a from an external source. This approach downloads the external data to a local machine. In this study, the specific dataset that will be used is the ERA5 dataset from the European Center for Medium-range Weather Forecast (ECMWF). Specifically, the ERA5 dataset is a reanalysis which means it combines observational data from all over the world in order to present a more complete picture of the weather system. This combination process, named 4D variational data assimilation, takes into account the physics known to be involved in the system. As an example weather station data which take measurements such as wind speed, humidity, and temperature of the immediate surroundings of the weather station is very limited in giving a broader picture of weather over a larger area. Even technologies like earth observations satellites are limited to what they can observe at any single point in time as most cannot see the entirety of the earth's surface at once. The ERA5 dataset specifically, assimilates previous forecasts with observational data every 12 hours. The ERA5 reanalysis is available at the following url (\url{https://cds.climate.copernicus.eu/datasets/reanalysis-era5-single-levels}).

The original dataset made available by ECMWF has a grid size of \ang{0.25} by \ang{0.25} in latitude and longitude for atmospheric data. Data of oceanic waves are also available at a coarser grid size of \ang{0.5} by \ang{0.5}. However, we will use a cloud optimized version called WeatherBench2 that is easier to retrieve \autocite{raspWeatherBench2Benchmark2023}. This is because the data is available in several more grid sizes such as \ang{1.5}. In addition, the data can be readily downloaded to a local machine without waiting for further processing on the server. The guide for working with the dataset is available at (\url{https://weatherbench2.readthedocs.io/en/latest/data-guide.html}). In our specific case, we use the version labeled \verb|1959-2023_01_10-6h-64x32_equiangular_conservative.zarr|. This dataset spans from year 1959 to 2023 with a grid size of \ang{5.625} or resolution of 64 by 32 which spans from \ang{-87.19} to \ang{87.19} in latitude and \ang{0.0} to \ang{354.4} in longitude. The dataset is also downsampled to 6 hour intervals that starts at midnight UTC January 1959 1\textsuperscript{st} and ends at 18:00 UTC January 10\textsuperscript{th} 2023.

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
  import xarray as xr
  xr.open_zarr(
      "DATASET_ADDRESS",
      storage_options = {"token": "anon"},
  )
  \end{lstlisting}
  \caption{Example of using Xarray to access a publicly accessible Zarr dataset hosted on Google Cloud Storage Bucket.} \label{fig:access_remote_zarr}
\end{figure}

The dataset includes 62 variables with some variables also spanning 13 discrete vertical levels in the atmosphere. For our use, we will only be using the 2-meter temperature of the atmosphere. This is the air temperature 2 meters above the surface be it land, sea, or inland water. This variable is in units of kelvin. The values of this variable is obtained by interpolating between the model values at the surface and the lowest model vertical level. We also limit the time range to .... An example of the data that was retrieved can be seen in .
% TODO: talk about selecting the time here
% ECMWF description
% - Role
% - legitimacy
% - Weatherbench2
% data description
% - data assimilation, maybe reference a general description
% - variables
% - dimensions/coordinates (time, spatial) and resolution in each
% - climatology

@@

The second use of a dataset depends on whether it can be interpreted as a mapping of a previous function to the next function. In other words, a series of functions related to the previous function as a sum of the previous function and a perturbation. A commonly used form of this are time series. Generally, a time series can be represented as \lccref{eq:time_step_function}. This function describes the evolution of function \(f\) as a sum of some initial state \(f^0\) and perturbation which is its rate of change \(\dv{f}{t}\) at a point in time close to the current time multiplied by the time step \(\delta t\). This is the first order Taylor polynomial. Adding more terms with higher order derivatives can be done to increase the accuracy. This describes the same process that is used in the finite differences method.
\begin{equation}
  f^{t+1} = f^{t} + \delta t \times \dv{f}{t} + \dots \label{eq:time_step_function}
\end{equation}
In this case the previous function and any other independent variables like the forcing term involved in the right-hand side of the equation is used to predict the difference between the future state and the current state. It would mean that each sample function in the data would need to at least have two dimensions. This is because one dimension represents the evolution of the function over time, while the other dimensions define the spatial domain where the function's values are computed at each time step. This condition is true for both the Burgers' equation dataset and the WeatherBench 2 collection of ERA5 datasets. First, the Burgers' equation dataset, while is also able to be used in the first way as a mapping between the forcing term and the solution, one of the traditional ways of solving it is using the method of lines. This entails solving the dimensions other that time with methods such as the finite differences method or the spectral method. The time dimension is then separately solved using the finite differences method such as the form shown in \lccref{eq:time_step_function} \autocite{schiesserNumericalMethodLines2012,sadikuSimpleIntroductionMethod2000}. This is the form that the Burgers' equation dataset will be used in. In this case, algebraic manipulation of \lccref{eq:forced_viscous_burgers} results in the time derivative of the solution as in \lccref{eq:forced_viscous_burgers_time_derivative}.
\begin{align}
  \pdv{u}{t} = f - u\pdv{u}{x} + \nu\pdv[2]{u}{x} \label{eq:forced_viscous_burgers_time_derivative}
\end{align}
The most important component of the right-hand side of the equation, the time derivative, is expressed in terms of the solution function, the forcing term and any other partial derivatives. Because of this, it means we can train the model to learn the relationship between the solution function combined with the forcing term and the time derivative. Putting this together with \lccref{eq:time_step_function}, it is possible to predict the solution function in a future time based on the current solution function and forcing term. The input functions in this case are both the value of the solution function and the forcing term at the current time step. The output function is the time derivative which is computed by rearranging \lccref{eq:time_step_function} resulting in \lccref{eq:time_derivative_time_step_function}.
\begin{equation}
  \dv{f}{t} = \frac{f^{t+1} - f^{t}}{\delta t} \label{eq:time_derivative_time_step_function}
\end{equation}
Perhaps in a more obvious manner, the same is also applied to the WeatherBench 2 collection of ERA5 datasets. The 2-meter temperature is a series of snapshots in time of the distribution of temperature all over the world at 2 meters above the surface. The input function in this case is the temperature distribution at the current time step. The output function is the time derivative of the temperature distribution computed using \lccref{eq:time_derivative_time_step_function}.

@@

\subsection{Model Evaluation}
\noindent The fourth phase of the computational model is the evaluation of the learned model. This phase starts with formatting the inputs. If the inputs aren't scaled yet, then it is scaled using the scaler fitted on the training set. This is the case for new data such as new measurements or an analytical solution of the learned PDE\@. Once the inputs that is to be evaluated is scaled in the same way as the training inputs, the model can predict the corresponding outputs.

@@

In addition, the software product is developed within certain limitations. The limitations are:
\begin{enumerate}
  \item The operating and implementation language of the library is Python.
  \item The library is to be used in user code
  \item The library depends on other libraries including the following core dependencies Pytorch, Numpy, Xarray, Zarr, Pandas, and tqdm.
\end{enumerate}

\subsection{Design} \label{sec:implementation_design}
\noindent In this subsection, the architectural design of the software product is explained. The architecture consists of modules with similar concerns. This was chosen because of the intuition built by the use of other libraries with a similar architecture. The broad aim of modeling PDEs with LSSVR and basis functions coefficients can be separated into smaller concerns of data, model, and basis functions. Each of these concerns are addressed by a corresponding module. The explanation of each module are as follows:
\begin{enumerate}
  \item Utilities

@@

\noindent The design in \lccref{sec:implementation_design} is implemented within the same iterative loop. As such, the feedback from tests is immediate, and any fixes can be implemented quickly. The implementation is done using the Python language and a number of Python libraries. To facilitate the management of dependencies, the implementation makes use of the package management tool Poetry \autocite{eustacePoetryPythonPackaging2024}. In addition, the Python version is managed using the tool pyenv. The implementation process starts with installation of the required dependencies. The dependencies are specified in the \verb|pyproject.toml| as shown in \lccref{fig:poetry_add_deps}.

@@

In the following, we will discuss the implementation of each module and how the phases of computational model relates to each implemented functionality. The base data structure that is used in the implementation is the Pytorch Tensor \autocite{Ansel_PyTorch_2_Faster_2024}. Tensors are a generalization of scalars and array-like structures such as vectors and matrices. This is essential since many computations that is done in the proposed model are based on matrices and higher dimensional arrays. Pytorch Tensors allow the elements to be of different data types such as floating point numbers of different precision or even complex numbers. In addition to basic operations for working with the Tensors such as matrix multiplications, Pytorch also provides many algorithms for use when working with Tensors such as linear equation solvers or the Fast Fourier Transform. Another reason for choosing to use Pytorch is the ability to parallelize computation using specialized processors such as GPUs. Lastly, Pytorch is widely used in the machine learning community which reflects on its reliability, wealth of community knowledge and support, and active development.

@@

\noindent The utilities module provide common tools that are used when working with the developed software. First, because the basis function used in this study is the Fourier basis, a large proportion of concern is the storage of Fourier basis coefficients. The Fourier series representation that is used in this study is the complex representation as show in \lccref{eq:fourier_series}. However, as previously mentioned, least squares support vector machines is constructed to work with real numbers. In order for LSSVR to learn the relationship between complex numbers, a representation of complex numbers in terms of real numbers is necessary. Since sample features and labels for LSSVR need to be flattened into one dimension which result in a matrix with rows of samples, the representation of complex numbers using real numbers can be done after features and labels have been flattened. An example of such a matrix of two samples with four features each can be seen in \lccref{eq:complex_matrix}. To represent the complex number features as real numbers, once can simply split a single complex number into two real numbers. When this is done to the entire matrix of complex numbers, the result is \lccref{eq:complex_matrix_real_rep}.
\begin{align}
  \begin{bmatrix}
    -1.03+0i & -0.52+0.39i & -0.54+0i & -0.52-0.39i \\
    -0.85+0i & -0.39+0.19i & 0.87+0i  & -0.39-0.19i
  \end{bmatrix}\label{eq:complex_matrix} \\
  \begin{bmatrix}
    -1.03 & 0 & -0.52 & 0.39 & -0.54 & 0 & -0.52 & -0.39 \\
    -0.85 & 0 & -0.39 & 0.19 & 0.87  & 0 & -0.39 & -0.19
  \end{bmatrix}\label{eq:complex_matrix_real_rep}
\end{align}
There are several ways to do this in Pytorch, such as extracting the real and imaginary element components into separate matrices and then interleaving the columns to create the real representation. However, a more performant version can be made by using built-in functions of the library that directly allows views of complex valued tensors as real tensors and vice versa. The \verb|view_as_real| and \verb|view_as_complex| functions allow the conversion between real and complex representations by representing complex valued tensors by adding a dimension of size two for each component of complex numbers. The resulting implementation can be seen in \lccref{fig:complex_matrix_conversion_implementation}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
def to_real_coeff(x: torch.Tensor) -> torch.Tensor:
    return torch.view_as_real(x).flatten(-2)

>>>>

% TODO: add figures and tables of generated data and some intermediate steps.
\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean_perturbed_solution.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_noise_levels}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_derivative_clean.pgf}
    \end{adjustbox}
    \caption{}\label{fig:derivative}
  \end{subfigure}
  \caption{(\subref{fig:antiderivative_noise_levels}) A generated function with no noise (clean), low noise level (5\%), medium noise level (10\%), and high noise levels (50\%).\ (\subref{fig:antiderivative}) Antiderivative function with no noise.\ (\subref{fig:derivative}) Derivative function with no noise.}
\end{figure}
% TODO: interesting note that model trains better on function to coefficient task when the antiderivative is computed from the randomly generated derivative than if the derivative is computed from the randomly generated antiderivative. THIS NEEDS A REASON AND ANALYSIS SO MIGHT NOT BE NEEDED IN THIS WRITING. MAYBE A GOOD IDEA FOR A SEPARATE PIECE

@@

After this, in the third step, the solution functions \(u \) are generated with 8 modes in time and 8 in space and 500 unique function samples for each viscosity value. Since we don't factor the viscosity in at this point, for ease of comparison, we ensure the same solutions are generated every time by using a new instance of the generator with the same seed number each time. This results in the solutions being the same for all viscosity and other parameters of the equation are adapted to this.

Next we account for the different viscosity values in this stage. The forcing term is computed using \lccref{eq:forced_viscous_burgers}. The solution field \(u\) and forcing term \(f\) are modeled with Fourier series in \lccrefs{eq:fourier_field,eq:fourier_force} where \(k\) is a vector of spatial and temporal wave number such that \(k_t\) is the temporal wave number and \(k_x\) is the spatial wave number.
\begin{align}
  u\left(x, t \right) & = \sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:fourier_field} \\
  f\left(x, t \right) & = \sum_{k_{x}}\sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:fourier_force}
\end{align}
Substituting the terms in \lccref{eq:forced_viscous_burgers} with the respective Fourier series, we get the equation \lccref{eq:forced_viscous_burgers_fourier_substitution}. A problem one notices is the nonlinear term \(u\pdv{u}{x}\). A naive approach would multiply all wave numbers terms with each other.
\begin{align}
  u\pdv{u}{x} = \left(\sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}\right) \times \left(\sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})\hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}\right)
\end{align}
This is computationally expensive operation with \(N^2\) multiplications \autocite{lariosMATH934BURGERS2021,robertsDealiasedConvolutionsPseudospectral2011,shenSpectralMethodsAlgorithms2011,orszagComparisonPseudospectralSpectral1972}. To avoid this complexity, the term is first reformulated into \(\pdv{x}(u^2/2)\). Then to avoid aliasing, the coefficients are padded with 50\% zeros such that the padded coefficients are 3/2 times the size of the original \autocite{orszagEliminationAliasingFiniteDifference1971, lariosMATH934BURGERS2021}. Then we use the padded coefficients to compute the function values in the physical domain using the inverse transform. And then, the point wise squaring operation is performed and transform the results back to spectral domain. The resulting coefficients are then trimmed back to their original size before padding. Finally, we multiply the resulting coefficients \(\hat{uu}_k\) by the derivative constants. This operation is much more efficient with only \(1.5 N\) multiplications. This is still more efficient than the naive approach even after taking into account the transforms involved adds \(O(2\times 1.5Nln(1.5N))\) operations if using a Fast Fourier Transform algorithm. Putting all the above together, results in \lccref{eq:forced_viscous_burgers_fourier_substitution}.
\begin{equation}
  \begin{split}
    \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x+k_{t}t)} =
     & \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}) \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}                                                                \\
     & + \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})\hat{uu}_k e^{2\pi i(k_{x}x+k_{t}t)}                                                              \\
     & - \nu\sum_{k_{x}} \sum_{k_{t}} {(2\pi i k_{x})}^2\hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:forced_viscous_burgers_fourier_substitution}
  \end{split}
\end{equation}
\begin{align}
  \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k & = \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}) \hat{u}_k + (2\pi i k_{x})\hat{uu}_k - \nu{(2\pi i k_{x})}^2\hat{u}_k \label{eq:forced_viscous_burgers_fourier} \\
  \hat{f}_k                           & = (2\pi i k_{t}) \hat{u}_k + (2\pi i k_{x})\hat{uu}_k - \nu{(2\pi i k_{x})}^2\hat{u}_k \label{eq:forced_viscous_burgers_coeff}
\end{align}
After simplifying the equation, we get the coefficients as in \lccref{eq:forced_viscous_burgers_coeff}. Using this equation, the exact forcing term corresponding to the randomly generated solution can be computed with relatively low cost. Finally, the function values of both the solutions and forcing terms are perturbed by adding Gaussian noise with a mean of zero and standard deviation of 10\% of the function value standard deviation to the values of the inverse transform. The perturbed coefficients are then recomputed from the sum and the final perturbed functions are obtained.

An example of the solution and forcing term pairs that has been generated is shown in \lccref{fig:burgers_data}. Using the same seed we can see that the solution functions of the same sample are the same. However, with the forcing terms, we see much larger amplitudes for higher frequencies for larger viscosity values. This is inline with \lccref{eq:forced_viscous_burgers_coeff}. The squared multiplier gets large much more quickly for higher frequencies due to the squaring operation. The since this effect is controlled by the viscosity via multiplication, the higher the viscosity, the larger the high frequency amplitudes of the forcing term. This, in other words, means that more viscous fluids require a stronger forcing term to affect the solution in the same way with as fluids with a lower viscosity.

@@

\section{Data Retrieval}\label{sec:data_retrieval}
\noindent The second data acquisition approach retrieves data of a from an external source. This approach downloads the external data to a local machine. In this study, the specific dataset that will be used is the ERA5 dataset from the European Center for Medium-range Weather Forecast (ECMWF). Specifically, the ERA5 dataset is a reanalysis which means it combines observational data from all over the world in order to present a more complete picture of the weather system. This combination process, named 4D variational data assimilation, takes into account the physics known to be involved in the system. As an example weather station data which take measurements such as wind speed, humidity, and temperature of the immediate surroundings of the weather station is very limited in giving a broader picture of weather over a larger area. Even technologies like earth observations satellites are limited to what they can observe at any single point in time as most cannot see the entirety of the earth's surface at once. The ERA5 dataset specifically, assimilates previous forecasts with observational data every 12 hours. The ERA5 reanalysis is available at the following URL (\url{https://cds.climate.copernicus.eu/datasets/reanalysis-era5-single-levels}).

The original dataset made available by ECMWF has a grid size of \ang{0.25} by \ang{0.25} in latitude and longitude for atmospheric data. Data of oceanic waves are also available at a coarser grid size of \ang{0.5} by \ang{0.5}. However, we will use a cloud optimized version called WeatherBench2 that is easier to retrieve \autocite{raspWeatherBench2Benchmark2023}. This is because the data is available in several more grid sizes such as \ang{1.5}. In addition, the data can be readily downloaded to a local machine without waiting for further processing on the server. The guide for working with the dataset is available at (\url{https://weatherbench2.readthedocs.io/en/latest/data-guide.html}). In our specific case, we use the version labeled \verb|1959-2023_01_10-6h-64x32_equiangular_conservative.zarr|. This dataset spans from year 1959 to 2023 with a grid size of \ang{5.625} or resolution of 64 by 32 which spans from \ang{-87.19} to \ang{87.19} in latitude and \ang{0.0} to \ang{354.4} in longitude. The dataset is also resampled down to 6-hour intervals that starts at midnight UTC January 1959 1\textsuperscript{st} and ends at 18:00 UTC January 10\textsuperscript{th} 2023.

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
  import xarray as xr
  xr.open_zarr(
      "DATASET_ADDRESS",
      storage_options = {"token": "anon"},
  )
  \end{lstlisting}
  \caption{Example of using Xarray to access a publicly accessible Zarr dataset hosted on Google Cloud Storage Bucket.}\label{fig:access_remote_zarr}
\end{figure}

The dataset includes 62 variables with some variables also spanning 13 discrete vertical levels in the atmosphere. For our use, we will only be using the 2-meter temperature of the atmosphere. This is the air temperature 2 meters above the surface be it land, sea, or inland water. This variable is in units of kelvin. The values of this variable is obtained by interpolating between the model values at the surface and the lowest model vertical level. We also limit the time range to ... An example of the data that was retrieved can be seen in .
% TODO: talk about selecting the time here
% ECMWF description
% - Role
% - legitimacy
% - Weatherbench2
% data description
% - data assimilation, maybe reference a general description
% - variables
% - dimensions/coordinates (time, spatial) and resolution in each
% - climatology

@@

The second use of a dataset depends on whether it can be interpreted as a mapping of a previous function to the next function. In other words, a series of functions related to the previous function as a sum of the previous function and a perturbation. A commonly used form of this are time series. Generally, a time series can be represented as \lccref{eq:time_step_function}. This function describes the evolution of function \(f\) as a sum of some initial state \(f^0\) and perturbation which is its rate of change \(\dv{f}{t}\) at a point in time close to the current time multiplied by the time step \(\delta t\). This is the first order Taylor polynomial. Adding more terms with higher order derivatives can be done to increase the accuracy. This describes the same process that is used in the finite difference method.
\begin{equation}
  f^{t+1} = f^{t} + \delta t \times \dv{f}{t} + \dots \label{eq:time_step_function}
\end{equation}
In this case the previous function and any other independent variables like the forcing term involved in the right-hand side of the equation is used to predict the difference between the future state and the current state. It would mean that each sample function in the data would need to at least have two dimensions. This is because one dimension represents the evolution of the function over time, while the other dimensions define the spatial domain where the function's values are computed at each time step. This condition is true for both the Burgers' equation dataset and the WeatherBench 2 collection of ERA5 datasets. First, the Burgers' equation dataset, while is also able to be used in the first way as a mapping between the forcing term and the solution, one of the traditional ways of solving it is using the method of lines. This entails solving the dimensions other that time with methods such as the finite differences method or the spectral method. The time dimension is then separately solved using the finite differences method such as the form shown in \lccref{eq:time_step_function} \autocite{schiesserNumericalMethodLines2012,sadikuSimpleIntroductionMethod2000}. This is the form that the Burgers' equation dataset will be used in. In this case, algebraic manipulation of \lccref{eq:forced_viscous_burgers} results in the time derivative of the solution as in \lccref{eq:forced_viscous_burgers_time_derivative}.
\begin{align}
  \pdv{u}{t} = f - u\pdv{u}{x} + \nu\pdv[2]{u}{x} \label{eq:forced_viscous_burgers_time_derivative}
\end{align}
The most important component of the right-hand side of the equation, the time derivative, is expressed in terms of the solution function, the forcing term and any other partial derivatives. Because of this, it means we can train the model to learn the relationship between the solution function combined with the forcing term and the time derivative. Putting this together with \lccref{eq:time_step_function}, it is possible to predict the solution function in a future time based on the current solution function and forcing term. The input functions in this case are both the value of the solution function and the forcing term at the current time step. The output function is the time derivative which is computed by rearranging \lccref{eq:time_step_function} resulting in \lccref{eq:time_derivative_time_step_function}.
\begin{equation}
  \dv{f}{t} = \frac{f^{t+1} - f^{t}}{\delta t} \label{eq:time_derivative_time_step_function}
\end{equation}
Perhaps in a more obvious manner, the same is also applied to the WeatherBench 2 collection of ERA5 datasets. The 2-meter temperature is a series of snapshots in time of the distribution of temperature all over the world at 2 meters above the surface. The input function in this case is the temperature distribution at the current time step. The output function is the time derivative of the temperature distribution computed using \lccref{eq:time_derivative_time_step_function}.

@@

\subsection{Model Evaluation}
\noindent The fourth phase of the computational model is the evaluation of the learned model. This phase starts with formatting the inputs. If the inputs aren't scaled yet, then it is scaled using the scaling function fitted on the training set. This is the case for new data such as new measurements or an analytical solution of the learned PDE\@. Once the inputs that is to be evaluated is scaled in the same way as the training inputs, the model can predict the corresponding outputs.

@@

In addition, the software product is developed within certain limitations. The limitations are:
\begin{enumerate}
  \item The operating and implementation language of the library is Python.
  \item The library is to be used in user code
  \item The library depends on other libraries including the following core dependencies PyTorch, NumPy, Xarray, Zarr, Pandas, and tqdm.
\end{enumerate}

\subsection{Design}\label{sec:implementation_design}
\noindent In this subsection, the architectural design of the software product is explained. The architecture consists of modules with similar concerns. This was chosen because of the intuition built by the use of other libraries with a similar architecture. The broad aim of modeling PDEs with LSSVR and basis functions coefficients can be separated into smaller concerns of data, model, and basis functions. Each of these concerns are addressed by a corresponding module. The explanation of each module are as follows:
\begin{enumerate}
  \item Utilities

@@

\noindent The design in \lccref{sec:implementation_design} is implemented within the same iterative loop. As such, the feedback from tests is immediate, and any fixes can be implemented quickly. The implementation is done using the Python language and a number of Python libraries. To facilitate the management of dependencies, the implementation makes use of the package management tool Poetry \autocite{eustacePoetryPythonPackaging2024}. In addition, the Python version is managed using the tool PyEnv. The implementation process starts with installation of the required dependencies. The dependencies are specified in the \verb|pyproject.toml| as shown in \lccref{fig:poetry_add_deps}.

@@

In the following, we will discuss the implementation of each module and how the phases of computational model relates to each implemented functionality. The base data structure that is used in the implementation is the PyTorch Tensor \autocite{Ansel_PyTorch_2_Faster_2024}. Tensors are a generalization of scalars and array-like structures such as vectors and matrices. This is essential since many computations that is done in the proposed model are based on matrices and higher dimensional arrays. PyTorch Tensors allow the elements to be of different data types such as floating point numbers of different precision or even complex numbers. In addition to basic operations for working with the Tensors such as matrix multiplications, PyTorch also provides many algorithms for use when working with Tensors such as linear equation solvers or the Fast Fourier Transform. Another reason for choosing to use PyTorch is the ability to parallelize computation using specialized processors such as GPUs. Lastly, PyTorch is widely used in the machine learning community which reflects on its reliability, wealth of community knowledge and support, and active development.

@@

\noindent The utilities module provide common tools that are used when working with the developed software. First, because the basis function used in this study is the Fourier basis, a large proportion of concern is the storage of Fourier basis coefficients. The Fourier series representation that is used in this study is the complex representation as show in \lccref{eq:fourier_series}. However, as previously mentioned, least squares support vector machines is constructed to work with real numbers. In order for LSSVR to learn the relationship between complex numbers, a representation of complex numbers in terms of real numbers is necessary. Since sample features and labels for LSSVR need to be flattened into one dimension which result in a matrix with rows of samples, the representation of complex numbers using real numbers can be done after features and labels have been flattened. An example of such a matrix of two samples with four features each can be seen in \lccref{eq:complex_matrix}. To represent the complex number features as real numbers, once can simply split a single complex number into two real numbers. When this is done to the entire matrix of complex numbers, the result is \lccref{eq:complex_matrix_real_rep}.
\begin{align}
  \begin{bmatrix}
    -1.03+0i & -0.52+0.39i & -0.54+0i & -0.52-0.39i \\
    -0.85+0i & -0.39+0.19i & 0.87+0i  & -0.39-0.19i
  \end{bmatrix}\label{eq:complex_matrix} \\
  \begin{bmatrix}
    -1.03 & 0 & -0.52 & 0.39 & -0.54 & 0 & -0.52 & -0.39 \\
    -0.85 & 0 & -0.39 & 0.19 & 0.87  & 0 & -0.39 & -0.19
  \end{bmatrix}\label{eq:complex_matrix_real_rep}
\end{align}
There are several ways to do this in PyTorch, such as extracting the real and imaginary element components into separate matrices and then interleaving the columns to create the real representation. However, a more performant version can be made by using built-in functions of the library that directly allows views of complex valued tensors as real tensors and vice versa. The \verb|view_as_real| and \verb|view_as_complex| functions allow the conversion between real and complex representations by representing complex valued tensors by adding a dimension of size two for each component of complex numbers. The resulting implementation can be seen in \lccref{fig:complex_matrix_conversion_implementation}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
def to_real_coeff(x: torch.Tensor) -> torch.Tensor:
    return torch.view_as_real(x).flatten(-2)
